---
title: "Advanced Features of tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Features of tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Advanced Features

This vignette covers the advanced features and configurations of the `tldrAI` package. These features are designed to enhance your experience with AI-powered R function documentation, providing more control, better performance, and additional customization options.

## In-depth Configuration Options

The `tldrAI` package offers comprehensive configuration options through the `tldr_config()` function. Here's a complete list of configuration parameters:

```{r}
tldr_config(
  # API Keys
  api_key = "your_claude_api_key",
  openai_api_key = "your_openai_api_key",
  
  # Provider settings
  provider = "claude",                # Default provider: "claude" or "openai"
  model = "claude-3-opus-20240229",   # Claude model
  openai_model = "gpt-4-turbo",       # OpenAI model
  
  # Fallback mechanism
  enable_fallback = TRUE,             # Enable fallback to another provider
  fallback_provider = "openai",       # Provider to use as fallback
  
  # Cache settings
  cache_enabled = TRUE,               # Enable/disable response caching
  cache_dir = "custom/cache/path",    # Custom cache directory
  cache_ttl = 30,                     # Cache time-to-live in days
  offline_mode = FALSE,               # Operate in offline mode (cache only)
  
  # Usage settings
  verbose_default = FALSE,            # Default verbosity setting
  examples_default = 2,               # Default number of examples
  character_voice = "wise_mentor",    # Default character voice
  
  # UI settings
  show_progress = TRUE,               # Show progress bar during API calls
  debug_mode = FALSE,                 # Show debug information
  
  # Performance settings
  async_mode = FALSE,                 # Enable asynchronous API calls
  timeout = 60,                       # API request timeout in seconds
  max_retries = 3,                    # Maximum number of retry attempts
  
  # Visualization settings
  visualize = TRUE,                   # Enable visualizations by default
  visualization_type = "diagram"      # Default visualization type
)
```

### Model Selection

You can choose from various LLM models with different capabilities and speed/cost tradeoffs:

```{r}
# Claude models
tldr_config(provider = "claude", model = "claude-3-opus-20240229")   # Most capable, slower
tldr_config(provider = "claude", model = "claude-3-sonnet-20240229") # Balanced performance
tldr_config(provider = "claude", model = "claude-3-haiku-20240307")  # Fastest, less capable

# OpenAI models
tldr_config(provider = "openai", openai_model = "gpt-4o")          # Latest model
tldr_config(provider = "openai", openai_model = "gpt-4-turbo")     # Balanced performance
tldr_config(provider = "openai", openai_model = "gpt-3.5-turbo")   # Faster, less capable
```

## Context Awareness with Privacy Settings

Contextual awareness allows `tldrAI` to analyze your R environment and provide tailored examples. You can configure exactly what data is analyzed and how it's handled:

```{r}
tldr_context_config(
  # Core settings
  enable_context_awareness = TRUE,    # Enable contextual awareness by default
  
  # Data sources to analyze
  analyze_data_frames = TRUE,         # Analyze data frames in the environment
  analyze_packages = TRUE,            # Analyze loaded packages
  analyze_history = TRUE,             # Analyze command history
  
  # Privacy controls
  anonymize_data = TRUE,              # Don't include actual data samples
  
  # Sampling limits
  max_rows_sample = 5,                # Maximum rows to sample
  max_cols_sample = 5,                # Maximum columns to sample
  
  # Information detail
  include_row_count = TRUE,           # Include row counts in context
  include_class_info = TRUE,          # Include class information
  include_column_types = TRUE,        # Include column types in context
  
  # History analysis
  max_history_commands = 10           # Maximum history commands to analyze
)
```

### Privacy and Data Protection

When `anonymize_data = TRUE`, the following privacy protections are applied:

1. **No actual data values** are sent to the LLM provider
2. Only **structural information** like column names, data types, and dimensions is included
3. Command history is filtered to remove potentially sensitive information

You can review exactly what would be shared with:

```{r}
tldr_test_context()
```

## Asynchronous API Usage

For improved workflow, you can make asynchronous API calls:

```{r}
# Make an asynchronous request
tldr("mean", async = TRUE)

# Continue working on other tasks...
x <- 1:100
summary(x)
hist(x)

# Check and retrieve the result when ready
result <- tldr_check_async()

# Check without waiting (returns NULL if not ready)
tldr_check_async(wait = FALSE)

# Wait with a custom timeout (in seconds)
tldr_check_async(timeout = 60)

# Enable async mode by default
tldr_config(async_mode = TRUE)
```

The asynchronous mode requires the `future` and `promises` packages:

```{r}
install.packages(c("future", "promises"))
```

## Performance Optimization

Optimize API performance with these settings:

```{r}
# Adjust timeout for slow connections
tldr_config(timeout = 120)  # 2 minutes

# Increase retry attempts for unreliable networks
tldr_config(max_retries = 5)

# Configure cache behavior
tldr_config(cache_ttl = 90)  # Keep cache valid for 90 days
```

### Provider Fallback

Configure automatic fallback to a secondary provider if the primary one fails:

```{r}
# Enable fallback from Claude to OpenAI
tldr_config(
  provider = "claude",
  enable_fallback = TRUE,
  fallback_provider = "openai"
)

# Or enable fallback from OpenAI to Claude
tldr_config(
  provider = "openai",
  enable_fallback = TRUE,
  fallback_provider = "claude"
)
```

## Advanced Visualization Options

`tldrAI` provides a powerful visualization system with multiple options:

### Visualization Types

```{r}
# Function diagram
tldr("mean", vis_type = "diagram")

# Logical flow diagram
tldr("if", vis_type = "flowchart")

# Data transformation flow diagram
tldr("dplyr::filter", vis_type = "data_flow")

# Function network diagram showing related functions
tldr("lm", vis_type = "function_network")

# Syntax highlighted code with interactive elements
tldr("mean", vis_type = "code_highlight")
```

### Visualization Configuration

Configure visualization settings:

```{r}
# Configure visualization settings
tldr_visualization_config(
  enable = TRUE,                   # Enable visualizations
  default_type = "diagram",        # Default visualization type
  auto_install = TRUE,             # Automatically install required packages
  auto_export = TRUE,              # Automatically export visualizations
  export_format = "svg",           # Default export format
  export_dir = "~/R/visualizations" # Export directory
)
```

### Exporting Visualizations Programmatically

```{r}
# Get function metadata and create visualization
metadata <- get_function_metadata("mean")
vis <- create_visualization("mean", metadata, "diagram")

# Export in different formats
export_visualization(vis, "mean_diagram.svg")
export_visualization(vis, "mean_diagram.png", format = "png", width = 800, height = 600)
export_visualization(vis, "mean_diagram.txt", format = "txt")
```

## Using Character Voices Effectively

Character voices add personality to responses. Here's how to use them effectively:

### Choosing the Right Voice for Different Tasks

Each voice has a distinct personality that may be better suited for different learning tasks:

```{r}
# For initial learning - enthusiastic and encouraging
tldr("ggplot", voice = "enthusiastic_explorer")

# For debugging or critical analysis
tldr("lm", voice = "cynical_detective")

# For deep understanding of complex functions
tldr("apply", voice = "wise_mentor")

# For creative or unexpected insights
tldr("lapply", voice = "eccentric_scientist")

# For humor and entertainment
tldr("mean", voice = "theatrical_villain")
```

### Creating Voice Combinations with Context Awareness

Combine voices with contextual awareness for personalized learning:

```{r}
# Personalized examples with enthusiastic voice
tldr("filter", context = TRUE, voice = "enthusiastic_explorer")

# Critical analysis of how a function applies to your data
tldr("summarize", context = TRUE, voice = "cynical_detective")
```

## Error Handling and Troubleshooting

`tldrAI` includes several features for troubleshooting:

### Debug Mode

Enable debug mode to see detailed information about what's happening:

```{r}
# Enable debug mode
tldr_debug(TRUE)

# Run a command to see debug information
tldr("mean")

# Disable debug mode
tldr_debug(FALSE)
```

### Testing API Connections

Test connections to providers:

```{r}
# Test connection to Claude
tldr_test_connection("claude")

# Test connection to OpenAI
tldr_test_connection("openai")
```

### Function Details

Get detailed information about a function:

```{r}
# Get function details
tldr_function_details("mean")
```

## Workflow Optimization

For the most efficient workflow:

### Set Up Defaults for Your Style

```{r}
# If you prefer verbose output
tldr_config(verbose_default = TRUE)

# If you like many examples
tldr_config(examples_default = 5)

# If you always want contextual examples
tldr_context_config(enable_context_awareness = TRUE)

# If you prefer a specific character voice
tldr_config(character_voice = "wise_mentor")

# If you always want visualizations
tldr_config(visualize = TRUE, visualization_type = "flowchart")
```

### Create Custom R Functions for Common Workflows

```{r}
# Quick tldr with your preferred settings
quick_tldr <- function(func_name) {
  tldr(func_name, verbose = TRUE, context = TRUE, voice = "enthusiastic_explorer")
}

# Visual tldr that always shows diagrams
visual_tldr <- function(func_name) {
  tldr(func_name, vis_type = "function_network", export_visualization = TRUE)
}
```

## Conclusion

The advanced features in `tldrAI` allow you to tailor the experience to your specific needs, whether you're focused on privacy, performance, visualization, or personalizing the learning experience. By combining these features, you can create a highly customized environment for exploring and understanding R functions.

For any issues or feature requests, you can open a GitHub issue through:

```{r}
tldr_open_repo()
```