---
title: "Secure API Key Management with tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Secure API Key Management with tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

API keys are credentials that provide access to cloud services. In the case of `tldrAI`, they're required to communicate with Large Language Model (LLM) providers like Anthropic's Claude and OpenAI. Properly managing these keys is crucial for security and convenience.

This vignette explains the secure API key management system in `tldrAI`, which offers multiple options for storing and retrieving API keys with a focus on security best practices.

## Why Secure API Key Management Matters

API keys for LLM providers like Claude and OpenAI:

1. **Grant access to paid services** that incur costs based on usage
2. **Can be misused** if obtained by unauthorized parties
3. **Should not be shared** in code repositories or public spaces
4. **Represent a personal account credential** that should be protected

The `tldrAI` package now offers secure storage of these API keys using your system's credential store via the `keyring` package, which is a significant security improvement over storing keys in plain text configuration files.

# Setting Up API Keys Securely

## Prerequisite: Install the keyring Package

The secure storage feature relies on the `keyring` package, which provides access to your operating system's secure credential store. If you haven't already installed it:

```{r}
install.packages("keyring")
```

## Securely Storing API Keys

The recommended way to store your API keys is using the `tldr_set_api_key()` function:

```{r}
library(tldrAI)

# Store Claude API key securely
tldr_set_api_key("your_claude_api_key", "claude")

# Store OpenAI API key securely
tldr_set_api_key("your_openai_api_key", "openai")
```

This function:

1. Stores your key in the system's secure credential store
2. Removes the key from the plain text configuration file if it was previously stored there
3. Uses service name "tldrAI" and username matching the provider ("claude" or "openai")

## Verifying API Key Storage

You can check if an API key is available for a specific provider:

```{r}
# Check if Claude API key is available
tldr_has_api_key("claude")

# Check if OpenAI API key is available
tldr_has_api_key("openai")
```

This function checks for keys in all possible locations (environment variables, keyring, and config file) without displaying the key values.

## Testing Your API Connection

To verify that your API keys are working correctly:

```{r}
# Test Claude connection
tldr_test_connection("claude")

# Test OpenAI connection
tldr_test_connection("openai")
```

# How tldrAI Manages API Keys

## Key Storage Hierarchy

The package uses a priority-based approach to find API keys:

1. **Environment variables** (`CLAUDE_API_KEY` or `OPENAI_API_KEY`) - highest priority
2. **System keyring** (secure credential store) - medium priority
3. **Configuration file** (plain text file) - lowest priority

This hierarchy ensures compatibility with different setups while prioritizing security:

- CI/CD environments and servers typically use environment variables
- Development machines benefit from secure keyring storage
- Legacy configurations still work without any changes

## Configuring Key Storage Behavior

You can control how `tldrAI` handles API keys with the `tldr_config()` function:

```{r}
# Enable secure keyring storage (default)
tldr_config(use_keyring = TRUE)

# Disable secure keyring storage (not recommended for regular use)
tldr_config(use_keyring = FALSE)
```

When `use_keyring = TRUE` (the default):

- The `tldr_config()` function will store new API keys in the system keyring if available
- You'll be prompted to migrate existing keys to the secure storage

# Migrating From Plain Text to Secure Storage

If you've been using `tldrAI` with API keys stored in the configuration file, you can easily migrate to secure storage.

## Automatic Migration Notification

When you load the package and it detects API keys in the configuration file, you'll see a notification suggesting migration:

```
tldrAI: Found API keys in configuration that could be stored more securely.
       Run tldr_key_migrate() to move keys to your system's secure keyring.
```

## Manual Migration

To migrate your existing keys:

```{r}
# Start the migration process
tldr_key_migrate()
```

This function:

1. Detects API keys stored in your configuration file
2. Prompts for confirmation before migration
3. Securely stores keys in the system keyring
4. Removes keys from the configuration file
5. Returns a report of which keys were migrated

## Clearing Stored API Keys

If you need to remove API keys from the secure storage:

```{r}
# Remove Claude API key
tldr_clear_api_key("claude")

# Remove OpenAI API key without confirmation
tldr_clear_api_key("openai", confirm = FALSE)
```

# Security Best Practices

## Do's and Don'ts

**Do:**

- ✓ Use the secure keyring storage when possible
- ✓ Use environment variables for CI/CD environments
- ✓ Regularly check your LLM provider usage dashboards
- ✓ Keep the `keyring` package updated
- ✓ Clear API keys when no longer needed

**Don't:**

- ✗ Share API keys in code, repositories, or public forums
- ✗ Store API keys in plain text files that might be backed up or synced
- ✗ Leave API keys in environment variables permanently on shared machines
- ✗ Ignore the migration prompts - secure storage is important

## Environment-Specific Recommendations

### Development Environment

```{r}
# Recommended setup for personal development
tldr_set_api_key("your_api_key", "claude")
tldr_config(use_keyring = TRUE)
```

### CI/CD Environment

```bash
# Set environment variables in your CI/CD system (not in R)
export CLAUDE_API_KEY=your_claude_api_key
export OPENAI_API_KEY=your_openai_api_key
```

The package will automatically find and use these environment variables.

### Server Environment

```{r}
# Option 1: Environment variables (preferred for servers)
# Set in your server environment, not in R code

# Option 2: If private user accounts need different keys
tldr_set_api_key("your_api_key", "claude")
```

# Technical Details

## How Secure Storage Works

The `keyring` package connects to your operating system's native credential store:

- **Windows**: Windows Credential Manager
- **macOS**: Keychain
- **Linux**: Secret Service API/libsecret (GNOME Keyring, KDE Wallet)

These systems:

1. Encrypt the stored credentials
2. Typically require authentication to access
3. Are designed specifically for securely storing sensitive information
4. Are not included in regular file backups or synchronization

## Fallback Mechanisms

If the `keyring` package is not available:

1. A warning message is displayed
2. The system falls back to using the configuration file
3. All functionality continues to work, just with reduced security

## Cross-Platform Compatibility

The secure storage system works across all major operating systems, but there are some platform-specific considerations:

- **Linux**: May require additional setup for the secret service (e.g., `libsecret-1-dev` package)
- **Headless systems**: May have limited keyring support, consider environment variables
- **RStudio Server/Cloud**: Varies by configuration, test the keyring availability

# Troubleshooting

## Common Issues and Solutions

**Problem**: "Failed to store API key in keyring" message

**Solutions**:
- Check if `keyring` package is installed: `install.packages("keyring")`
- On Linux, install required dependencies: `sudo apt-get install libsecret-1-dev`
- Verify your system's credential store is working properly

**Problem**: API key works in one session but not another

**Solutions**:
- Check if you're using environment variables that might be session-specific
- Ensure the keyring service is running on your system
- Verify permissions on the configuration directory

**Problem**: Need to use different API keys for different projects

**Solutions**:
- Use environment variables with project-specific settings
- Use different R profiles for different projects
- Set the API key at the start of each session

## Diagnosing Keyring Issues

If you're experiencing problems with the keyring integration:

```{r}
# Check if keyring package can be loaded
requireNamespace("keyring", quietly = TRUE)

# Test basic keyring functionality (if available)
if(requireNamespace("keyring", quietly = TRUE)) {
  tryCatch({
    keyring::key_set_with_value("test", "test", "test")
    keyring::key_get("test", "test")
    keyring::key_delete("test", "test")
    message("Keyring test successful")
  }, error = function(e) {
    message("Keyring test failed: ", e$message)
  })
}
```

# Conclusion

Secure API key management is an important aspect of working with LLM providers. The `tldrAI` package now offers a robust system for managing your API keys securely across different environments.

By using the secure storage options, you can:

- Keep your API keys protected from unauthorized access
- Avoid accidentally sharing credentials in repositories or backups
- Ensure your keys are available when needed
- Maintain compatibility with CI/CD and server environments

We recommend migrating to the secure storage system if you're currently using configuration files for your API keys.