---
title: "Error Analysis with tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Error Analysis with tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

R error messages can often be cryptic and difficult to understand, especially for beginners or when working with complex packages. The `tldr_error()` function in the `tldrAI` package solves this problem by providing AI-powered explanations of error messages, practical suggestions for fixing them, and working code examples.

This vignette explains how to use the error analysis capabilities of `tldrAI` to quickly understand and resolve R errors.

## Why AI-Powered Error Analysis?

Traditional error messages in R provide information about what went wrong, but often lack:

1. **Clear explanations** in user-friendly language
2. **Context-aware solutions** that consider your specific environment
3. **Working examples** that demonstrate how to fix the issue
4. **Educational insights** that help you avoid similar errors in the future

The `tldr_error()` function addresses these limitations by leveraging Large Language Models (LLMs) to provide comprehensive, human-readable error analysis.

# Basic Usage

## Analyzing a Simple Error Message

The most straightforward way to use `tldr_error()` is to provide an error message as a string:

```{r}
library(tldrAI)

# Analyze a simple error message
tldr_error("Error in mean(x) : object 'x' not found")
```

This will generate a comprehensive explanation of what the error means, why it occurred, and how to fix it.

## Analyzing an Error Object

You can also analyze error objects directly from `tryCatch()`:

```{r}
# Use with tryCatch to analyze errors in real-time
result <- tryCatch(
  {
    # Code that might cause an error
    mean(nonexistent_variable)
  },
  error = function(e) {
    # Analyze the error
    tldr_error(e)
    
    # Return NULL or some default value
    NULL
  }
)
```

## Automatically Capturing the Last Error

If you've just encountered an error in your R session, you can analyze it without having to retype it:

```{r}
# This will cause an error
mean(nonexistent_variable)

# Analyze the error that just occurred
tldr_error(auto_capture = TRUE)
```

# Customizing Error Analysis

## Including Code Context

For the most accurate analysis, you can provide the code that caused the error:

```{r}
# Provide both the error and the code
tldr_error(
  "Error: could not find function '%>%'",
  code = "data %>% filter(x > 5)"
)
```

This helps the AI understand exactly what you were trying to do, so it can provide more targeted advice.

## Adjusting Detail Level

You can control the level of detail in the recommendations and the number of examples:

```{r}
# Get more examples
tldr_error(
  "Error in library(nonexistent) : there is no package called 'nonexistent'",
  examples = 3
)

# Disable recommendations
tldr_error(
  "Error in read.csv(file) : object 'file' not found",
  recommend = FALSE
)
```

## Using Character Voices

Like other `tldrAI` functions, you can apply character voices to make error explanations more engaging:

```{r}
# Get error explanation with a professor voice
tldr_error(
  "Error in if (x > 0) { : missing value where TRUE/FALSE needed",
  voice = "professor"
)

# List available voices
tldr_list_voices()
```

# Context-Aware Error Analysis

## Leveraging Your R Environment

By default, `tldr_error()` analyzes your current R environment to provide more relevant explanations:

```{r}
# Create some data
df <- data.frame(a = 1:5, b = 6:10)

# Try a function that will error
ggplot(df, aes(x = a, y = b)) + geom_point()

# Analyze with context awareness
tldr_error(auto_capture = TRUE)
```

With context awareness enabled, the function can:

- Recognize loaded packages and suggest missing ones
- Reference your actual data frames and variables
- Understand your recent command history
- Provide examples that work with your specific environment

You can disable context awareness if you prefer more general explanations:

```{r}
tldr_error(
  "Error in ggplot(df, aes(x = a, y = b)) : could not find function 'ggplot'",
  context = FALSE
)
```

# Automatic Error Handling

## Auto-Explaining All Errors

For a more interactive experience, you can enable automatic error explanations:

```{r}
# Enable automatic error explanations
tldr_auto_error(TRUE)

# Now any error will be automatically explained
mean(nonexistent_variable)
```

This replaces R's default error handler with one that automatically analyzes and explains errors as they occur. You can disable this feature when you no longer need it:

```{r}
# Disable automatic error explanations
tldr_auto_error(FALSE)
```

# Error Patterns and Solutions

The `tldr_error()` function recognizes many common error patterns in R and provides targeted solutions:

## Missing Objects

```{r}
tldr_error("Error: object 'df' not found")
```

This will explain why objects can be missing and suggest solutions like checking for typos, properly loading data, or ensuring the object was actually created.

## Function Not Found

```{r}
tldr_error("Error: could not find function 'ggplot'")
```

The explanation will identify likely causes like missing packages, namespace issues, or typos, and provide appropriate installation commands or namespace solutions.

## Syntax Errors

```{r}
tldr_error("Error: unexpected '}' in \"if(x > 0) {}\"")
```

Syntax errors are explained with clear guidance on R's syntax rules and examples of correct structure.

## Type Mismatches

```{r}
tldr_error("Error: non-numeric argument to binary operator")
```

The function will explain type compatibility issues and how to properly convert between types.

## Package-Specific Errors

```{r}
tldr_error("Error: geom_point requires the following missing aesthetics: x")
```

For common packages like ggplot2, dplyr, and others, you'll get package-specific advice that explains the error in the context of how that package works.

# Advanced Usage

## Integration with Development Workflows

You can integrate error analysis into your development workflows:

```{r}
# In testing scripts
run_tests <- function() {
  tryCatch(
    {
      # Your test code here
      test_results <- run_your_tests()
      test_results
    },
    error = function(e) {
      cat("Test failed with error:\n")
      tldr_error(e, context = TRUE)
      NULL
    }
  )
}
```

## Debugging Complex Errors

For complex errors, `tldr_error()` can analyze the traceback to identify the root cause:

```{r}
complex_function <- function() {
  inner_function()
}

inner_function <- function() {
  deepest_function()
}

deepest_function <- function() {
  # This will error
  1 + "string"
}

# Capture and analyze complex error
tryCatch(
  complex_function(),
  error = function(e) {
    tldr_error(e)
  }
)
```

# Conclusion

The `tldr_error()` function transforms frustrating R error messages into learning opportunities. By providing clear explanations, practical solutions, and working examples, it helps you resolve issues quickly and deepen your understanding of R.

We recommend using this feature when:

- Learning R or a new package
- Debugging complex code
- Teaching or mentoring others
- Working with unfamiliar codebases

The AI-powered error analysis in `tldrAI` helps make the debugging process more efficient and educational, turning error messages from roadblocks into stepping stones for improvement.