---
title: "Getting Started with tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

`tldrAI` is an R package that provides AI-powered quick reference for R functions. It uses Large Language Models (LLMs) like Claude and OpenAI to generate concise, practical help text focused on examples and real-world usage. The goal is to maintain the brevity and example-focused philosophy of "tldr" guides while adding AI intelligence for highly relevant, personalized help.

Unlike traditional R help documentation, which can be verbose and focused on theoretical explanations, `tldrAI` focuses on:

- **Concise explanations**: Clear, straightforward descriptions of what functions do
- **Practical examples**: Real-world code examples showing common usage patterns
- **Contextual awareness**: Examples tailored to your actual data and environment
- **Personality**: Optional character voices to make learning more engaging

This vignette will guide you through setting up and using `tldrAI` for everyday R development.

## Installation

You can install the development version of `tldrAI` from GitHub with:

```{r}
# install.packages("devtools")
devtools::install_github("shandley/tldrAI")
```

## Getting Started

First, load the package:

```{r}
library(tldrAI)
```

### API Keys

`tldrAI` requires at least one API key to communicate with LLM providers. You can use either Claude (Anthropic), OpenAI, or both:

#### Setting up Claude API Key (Recommended)

1. Get a Claude API key from [Anthropic's console](https://console.anthropic.com/)
2. Configure tldrAI to use your key:

```{r}
tldr_config(api_key = "your_claude_api_key")
```

#### Setting up OpenAI API Key (Alternative)

1. Get an OpenAI API key from [OpenAI's platform](https://platform.openai.com/api-keys)
2. Configure tldrAI to use your key:

```{r}
tldr_config(openai_api_key = "your_openai_api_key")
```

You can test your API connection with:

```{r}
# Test Claude connection
tldr_test_connection("claude")

# Test OpenAI connection
tldr_test_connection("openai")
```

## Basic Usage

The main function is `tldr()`, which takes an R function name and returns concise help:

```{r}
# Get help for the mean function
tldr("mean")

# Get help for a function in a specific package
tldr("dplyr::filter")
```

### Adding More Detail

You can customize the level of detail returned:

```{r}
# Get more detailed help
tldr("mean", verbose = TRUE)

# Get more examples
tldr("mean", examples = 5)
```

### Using Different Providers

You can choose which LLM provider to use:

```{r}
# Use Claude by default (if configured)
tldr("mean")

# Explicitly use OpenAI
tldr("mean", provider = "openai")
```

## Contextual Awareness

One of the most powerful features of `tldrAI` is contextual awareness, which provides examples tailored to your actual environment:

```{r}
# Load some example data
library(dplyr)
data(mtcars)

# Get help with context
tldr("filter", context = TRUE)
```

The contextual help will include examples using your actual data frames, like `mtcars` in this example.

### Enabling Contextual Awareness by Default

You can enable contextual awareness for all queries:

```{r}
tldr_context_config(enable_context_awareness = TRUE)
```

### Privacy Settings

Configure what data is analyzed and shared with the AI:

```{r}
tldr_context_config(
  analyze_data_frames = TRUE,    # Analyze data frames in environment
  analyze_packages = TRUE,       # Look at loaded packages
  analyze_history = TRUE,        # Look at command history
  anonymize_data = TRUE,         # Don't include actual data values
  max_rows_sample = 3,           # Maximum rows to sample
  max_cols_sample = 5            # Maximum columns to sample
)
```

### Testing Context Analysis

You can see what information would be collected:

```{r}
tldr_test_context()
```

## Character Voices

Add personality to responses with character voices:

```{r}
# List available voices
tldr_list_voices()

# Try different voices
tldr("mean", voice = "enthusiastic_explorer")
tldr("sd", voice = "cynical_detective")
tldr("median", voice = "wise_mentor")
tldr("var", voice = "eccentric_scientist")

# Set a default voice
tldr_config(character_voice = "enthusiastic_explorer")
```

## Visualizations

`tldrAI` can generate visualizations to help understand functions better:

```{r}
# Get function visualization
tldr("mean", visualize = TRUE)

# Try different visualization types
tldr("mean", vis_type = "diagram")           # Simple function diagram
tldr("if", vis_type = "flowchart")           # Logical flow diagram
tldr("dplyr::filter", vis_type = "data_flow") # Data transformation flow
tldr("lm", vis_type = "function_network")     # Related functions network
tldr("mean", vis_type = "code_highlight")     # Syntax highlighted code
```

### Exporting Visualizations

You can export visualizations to files:

```{r}
# Export as SVG (default)
tldr("mean", visualize = TRUE, export_visualization = TRUE, 
     export_path = "mean_diagram.svg")

# Export as PNG
tldr("if", vis_type = "flowchart", export_visualization = TRUE, 
     export_path = "if_flowchart.png", export_format = "png")
```

## Offline Mode

`tldrAI` can work offline using cached responses:

```{r}
# Enable offline mode
tldr_offline(TRUE)

# Use cached responses
tldr("mean")

# Disable offline mode when back online
tldr_offline(FALSE)
```

## Cache Management

`tldrAI` automatically caches responses to improve performance and reduce API costs:

```{r}
# Clear all cached responses
tldr_cache_clear()

# Clear only expired cache entries
tldr_cache_clear(expired_only = TRUE)

# Set cache time-to-live (in days)
tldr_config(cache_ttl = 30)
```

## Configuration

You can configure various aspects of `tldrAI`:

```{r}
tldr_config(
  # Provider settings
  provider = "claude",                # Default provider
  enable_fallback = TRUE,             # Enable fallback to another provider
  fallback_provider = "openai",       # Provider to use as fallback
  
  # Usage settings
  verbose_default = FALSE,            # Default verbosity
  examples_default = 2,               # Default number of examples
  
  # UI settings
  show_progress = TRUE,               # Show progress bar during API calls
  
  # Cache settings
  cache_enabled = TRUE,               # Enable response caching
  cache_ttl = 30,                     # Cache time-to-live in days
  
  # Visualization settings
  visualize = TRUE,                   # Enable visualizations by default
  visualization_type = "diagram"      # Default visualization type
)
```

## Conclusion

`tldrAI` transforms how you get help in R, making it more concise, relevant, and personalized. It's especially helpful for:

- Learning new functions quickly
- Understanding complex functions with practical examples
- Getting examples tailored to your actual data
- Making R documentation more engaging with character voices

For more advanced features, check out the "Advanced Features" vignette.