---
title: "Package Analysis with tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package Analysis with tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

R has thousands of packages available on CRAN and other repositories, each with its own set of functions, datasets, and workflows. Learning a new package can be time-consuming, requiring you to read documentation, browse vignettes, and piece together how different functions work together.

The `tldr_package()` function in the `tldrAI` package solves this problem by providing AI-powered package analysis that quickly gets you up to speed with any R package. Rather than spending hours exploring package documentation, you can get a comprehensive overview in seconds.

This vignette explains how to use the package analysis capabilities of `tldrAI` to understand and work with R packages more efficiently.

## Why AI-Powered Package Analysis?

Traditional package documentation provides detailed information but often lacks:

1. **Clear organization** by task or workflow
2. **Relationships between functions** and how they work together
3. **Practical workflow examples** that combine multiple functions
4. **Function categorization** by purpose
5. **Concise overviews** that focus on the most important features

The `tldr_package()` function addresses these limitations by leveraging Large Language Models (LLMs) to analyze package structure, functions, datasets, and documentation, then synthesize this information into a clear, practical overview.

# Basic Usage

## Analyzing a Package

The most straightforward way to use `tldr_package()` is to provide a package name:

```{r}
library(tldrAI)

# Analyze the dplyr package
tldr_package("dplyr")
```

This will generate a comprehensive overview of the package, including its purpose, key functions, installation instructions, usage examples, and best practices.

## Package Must Be Installed

For best results, the package should be installed on your system. If it's not installed, `tldr_package()` will provide information about what the package likely does and installation instructions:

```{r}
# For a package that's not installed
tldr_package("packageNotInstalled")
```

# Customizing Package Analysis

## Focusing on Specific Aspects

You can focus the analysis on specific aspects of a package using the `focus` parameter:

```{r}
# Focus on functions
tldr_package("ggplot2", focus = "functions")

# Focus on datasets
tldr_package("datasets", focus = "datasets")

# Focus on vignettes
tldr_package("knitr", focus = "vignettes")

# Focus on dependencies
tldr_package("tidyr", focus = "dependencies")
```

The available focus options are:

- `"overview"`: General package purpose and key functions (default)
- `"functions"`: Detailed exploration of all exported functions
- `"datasets"`: Description of included datasets
- `"vignettes"`: Summary of package vignettes
- `"dependencies"`: Analysis of package dependencies and their purposes

## Adjusting Detail Level

Control the amount of detail in the analysis with the `detail` parameter:

```{r}
# Brief overview
tldr_package("lubridate", detail = "brief")

# Standard overview (default)
tldr_package("purrr", detail = "standard")

# Comprehensive analysis
tldr_package("dplyr", detail = "comprehensive")
```

The available detail levels are:

- `"brief"`: Quick overview with key functions only
- `"standard"`: Balanced detail with main function categories (default)
- `"comprehensive"`: Full details on all exported functions and features

## Workflow Examples

You can request more workflow examples to understand how to use the package in practice:

```{r}
# Request 5 workflow examples
tldr_package("tidyr", examples = 5)
```

## Function Categorization

By default, functions are grouped by purpose (data manipulation, visualization, etc.). You can disable this behavior if you prefer a flat list:

```{r}
# Disable function categorization
tldr_package("dplyr", categories = FALSE)
```

# Interactive Mode

The `interactive` parameter creates a simple console-based explorer that allows you to navigate through different sections of the package analysis:

```{r}
# Interactive exploration
tldr_package("shiny", interactive = TRUE)
```

In interactive mode, you can:

- Use arrow keys or numbers to navigate between sections
- Press Enter to view a selected section
- Press 'q' to quit and return to the R console

This mode is particularly useful for larger packages with many functions, where the full analysis might be overwhelming in a single view.

# Output Formats

You can control the output format using the `output` parameter:

```{r}
# Default console output with formatting
tldr_package("readr")

# Raw markdown output
tldr_package("readr", output = "markdown")

# HTML output (requires rmarkdown package)
tldr_package("readr", output = "html")
```

The available output formats are:

- `"console"`: Formatted console output (default)
- `"markdown"`: Raw markdown text
- `"html"`: HTML format (requires rmarkdown package)

The HTML output is particularly useful for sharing package analyses with others or including them in reports and presentations.

# Provider and Voice Options

Like other `tldrAI` functions, you can use different LLM providers and character voices:

```{r}
# Using a specific voice
tldr_package("tibble", voice = "enthusiastic_explorer")

# Using a specific provider
tldr_package("stringr", provider = "openai")
```

# Example: Learning the dplyr Package

Here's a practical example of how `tldr_package()` can help you learn the popular dplyr package:

```{r}
# Get a comprehensive overview of dplyr
tldr_package("dplyr", detail = "comprehensive", examples = 3)
```

This will provide:

1. A clear explanation of dplyr's purpose in the tidyverse ecosystem
2. Key functions grouped by tasks like filtering, summarizing, and joining
3. Multiple workflow examples showing how to chain operations with the pipe operator
4. Best practices for efficient data manipulation
5. Common patterns and anti-patterns

# Integration with Other tldrAI Features

The `tldr_package()` function is designed to work seamlessly with other tldrAI features:

```{r}
# First get an overview of the package
tldr_package("ggplot2")

# Then explore a specific function
tldr("ggplot2::ggplot")

# Analyze error messages related to the package
# This will error
ggplot(mtcars)
# Then analyze the error
tldr_error(auto_capture = TRUE)

# Explain code using this package
code <- "
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  geom_smooth(method = 'lm')
"
tldr_explain(code)
```

# Conclusion

The `tldr_package()` function transforms how you learn and use R packages by providing clear, practical overviews that focus on what matters most. It helps you:

- Quickly understand a package's purpose and capabilities
- Identify the most important functions and how they relate
- Learn from practical workflow examples
- Navigate through even the most complex packages

By making package documentation more accessible and user-friendly, `tldr_package()` helps you become more productive with new packages in minutes rather than hours or days.

Next time you're learning a new package or need a refresher on one you haven't used in a while, try `tldr_package()` to get up to speed quickly.