---
title: "Script Summarization with tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Script Summarization with tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Introduction

Understanding R scripts—both your own and those written by others—can be challenging. Scripts often contain complex data manipulations, statistical analyses, and visualizations without extensive documentation. The `tldr_summarize()` function in the `tldrAI` package leverages Large Language Models (LLMs) to generate intelligent, human-readable summaries of R scripts.

This vignette explains how to use the `tldr_summarize()` function to analyze R scripts, understand their purpose, structure, and functionality, and even generate visualizations of their components and relationships.

## Key Benefits of Script Summarization

Script summarization with `tldrAI` offers several advantages:

1. **Quick comprehension** of unfamiliar scripts without reading every line
2. **Identification of key components** like functions, data operations, and workflows
3. **Visualization options** for understanding dependencies, flow, and structure
4. **Customizable focus** to highlight specific aspects of the script
5. **Integration with context** from your current R environment
6. **Time savings** when onboarding to new projects or revisiting old code

# Basic Usage

## Summarizing a Single Script

The simplest way to use `tldr_summarize()` is to provide the path to an R script:

```{r}
library(tldrAI)

# Generate a summary of a single script
summary <- tldr_summarize("path/to/your/script.R")

# The result is displayed in the console by default
summary
```

The function analyzes the script and returns a structured summary with sections covering:

- Purpose and overview
- Key components and functionality
- Package dependencies
- Data flow and transformations
- Execution flow

## Summarizing Multiple Scripts

You can also analyze multiple related scripts at once:

```{r}
# Analyze multiple scripts together
summaries <- tldr_summarize(c(
  "data_prep.R",
  "analysis.R",
  "visualization.R"
))
```

When multiple scripts are provided, the function analyzes both individual scripts and their relationships, helping you understand how they work together.

# Customizing the Summary

## Focusing on Specific Aspects

The `focus` parameter allows you to emphasize different aspects of the script:

```{r}
# Focus on function definitions and relationships
tldr_summarize("utilities.R", focus = "functions")

# Focus on data transformations and flow
tldr_summarize("data_pipeline.R", focus = "data_flow")

# Focus on package dependencies
tldr_summarize("analysis.R", focus = "packages")

# Focus on execution workflow
tldr_summarize("processing.R", focus = "workflow")

# General overview (default)
tldr_summarize("script.R", focus = "general")
```

Each focus option tailors the summary to highlight different aspects of the script:

| Focus Option | Emphasizes |
|-------------|------------|
| "general" | Overall script purpose and functionality |
| "functions" | Functions defined, their purposes and relationships |
| "data_flow" | Data imports, transformations, and outputs |
| "packages" | Package dependencies and their usage |
| "workflow" | Execution flow and computational steps |

## Controlling Detail Level

The `detail` parameter lets you control the amount of information included:

```{r}
# Brief summary with essential information
tldr_summarize("script.R", detail = "brief")

# Standard detail (default)
tldr_summarize("script.R", detail = "standard")

# Comprehensive analysis with in-depth information
tldr_summarize("script.R", detail = "comprehensive")
```

# Visualizing Script Structure

One of the most powerful features of `tldr_summarize()` is the ability to generate visualizations of the script's structure.

## Enabling Visualizations

To include a visualization with your summary:

```{r}
# Generate a summary with visualization
summary_with_vis <- tldr_summarize("script.R", visualize = TRUE)

# The visualization is included in the output
summary_with_vis
```

## Visualization Types

The `vis_type` parameter allows you to choose from four different visualization types:

```{r}
# Function dependency visualization (default)
tldr_summarize("script.R", visualize = TRUE, vis_type = "dependency")

# Execution flow visualization
tldr_summarize("script.R", visualize = TRUE, vis_type = "flow")

# Package dependency visualization
tldr_summarize("script.R", visualize = TRUE, vis_type = "package")

# Script structure visualization
tldr_summarize("script.R", visualize = TRUE, vis_type = "structure")
```

### Dependency Visualization

The "dependency" visualization shows how functions within the script(s) depend on each other and on external packages:

```{r}
# Example of dependency visualization
dep_viz <- tldr_summarize("analysis.R", 
                         visualize = TRUE, 
                         vis_type = "dependency")
```

This visualization helps you understand:
- Which functions call which other functions
- Which packages are used by different functions
- How complex the function dependencies are

### Flow Visualization

The "flow" visualization represents the execution flow of the script:

```{r}
# Example of flow visualization
flow_viz <- tldr_summarize("data_pipeline.R", 
                          visualize = TRUE, 
                          vis_type = "flow")
```

This helps you understand:
- The overall sequence of operations
- Major code blocks and their purposes
- Data input and output points

### Package Visualization

The "package" visualization focuses on package dependencies:

```{r}
# Example of package visualization
pkg_viz <- tldr_summarize("analysis.R", 
                         visualize = TRUE, 
                         vis_type = "package")
```

This visualization helps you understand:
- Which packages the script depends on
- How packages are grouped by functionality
- The extent of package usage

### Structure Visualization

The "structure" visualization shows the overall structure of the script:

```{r}
# Example of structure visualization
struct_viz <- tldr_summarize("utilities.R", 
                            visualize = TRUE, 
                            vis_type = "structure")
```

This helps you understand:
- Functions defined in the script
- Global variables and their relationships
- Major code sections and their purposes

## Fallback Visualizations

If the required packages for generating graphical visualizations (`DiagrammeR` and/or `visNetwork`) are not available, `tldr_summarize()` will automatically fall back to ASCII-based visualizations that can be displayed in any console.

# Advanced Features

## Context-Aware Summaries

The `tldr_summarize()` function can incorporate context from your current R environment to provide more relevant summaries:

```{r}
# Load some packages and data
library(dplyr)
library(ggplot2)
data(mtcars)

# Generate a context-aware summary
tldr_summarize("visualization.R", context = TRUE)
```

With `context = TRUE`, the summary will consider:
- Currently loaded packages
- Available data frames
- Recent R commands
- Environment variables

This contextual information helps the summary be more relevant to your current work.

## Character Voices

Like other `tldrAI` functions, `tldr_summarize()` supports character voices to add personality to summaries:

```{r}
# List available voices
tldr_list_voices()

# Use a specific voice for the summary
tldr_summarize("script.R", voice = "professor")
```

Each voice provides a different style of communication while maintaining accuracy.

## LLM Provider Selection

You can choose which LLM provider to use for generating summaries:

```{r}
# Use Claude for the summary
tldr_summarize("script.R", provider = "claude")

# Use OpenAI for the summary
tldr_summarize("script.R", provider = "openai")
```

Different providers may have different strengths and capabilities.

## Caching and Refresh

Summaries are cached to improve performance and reduce API calls:

```{r}
# Use cached result if available (default)
tldr_summarize("script.R")

# Force a fresh summary
tldr_summarize("script.R", refresh = TRUE)
```

Cached results are stored based on the script's content, focus, detail level, and other parameters.

## Output Format

You can control the output format of the summary:

```{r}
# Console output with colored formatting (default)
tldr_summarize("script.R", output = "console")

# Markdown output for inclusion in documents
tldr_summarize("script.R", output = "markdown")
```

The markdown output is particularly useful for including summaries in R Markdown documents or reports.

# Practical Examples

## Analyzing a Data Preparation Script

```{r}
# Analysis of a data preparation script
tldr_summarize("data_prep.R", 
              focus = "data_flow", 
              visualize = TRUE, 
              vis_type = "flow")
```

This would provide a summary focusing on the data transformations with a flow visualization showing how data moves through the script.

## Understanding a Complex Analysis Script

```{r}
# Comprehensive analysis of a complex script
tldr_summarize("statistical_analysis.R", 
              detail = "comprehensive", 
              visualize = TRUE, 
              vis_type = "structure")
```

This provides an in-depth analysis with a structural visualization to help understand the components.

## Exploring Package Dependencies

```{r}
# Analysis focusing on package dependencies
tldr_summarize("visualization_functions.R", 
              focus = "packages", 
              visualize = TRUE, 
              vis_type = "package")
```

This summary would focus on which packages are used and why, with a visualization of package relationships.

## Documentation Generation

```{r}
# Generate documentation-style summary
markdown_summary <- tldr_summarize("helper_functions.R", 
                                  focus = "functions", 
                                  detail = "comprehensive", 
                                  output = "markdown")

# Could be saved to a file
cat(markdown_summary$summary, file = "helper_functions_documentation.md")
```

This approach can be used to generate documentation for script libraries.

# Technical Details

## Script Analysis Process

When you call `tldr_summarize()`, the function:

1. Validates input parameters and file paths
2. Extracts script content and metadata
3. Analyzes code structure:
   - Identifies package dependencies (`library()`, `require()`, and namespace (`::`) calls)
   - Detects function definitions and relationships
   - Identifies variables and data operations
   - Analyzes code blocks and execution flow
4. Constructs an appropriate prompt for the LLM
5. Sends the prompt to the selected LLM provider
6. Processes the response into a structured summary
7. Generates visualizations if requested
8. Returns and displays the results

## Return Value

The function returns a list containing:

```{r}
result <- tldr_summarize("script.R", visualize = TRUE)

# Components of the return value
names(result)
# [1] "summary"      "files"        "focus"        "provider"     "visualization"
```

- `summary`: The text summary (formatted according to the `output` parameter)
- `files`: The file paths that were analyzed
- `focus`: The focus used for the summary
- `provider`: The LLM provider that generated the summary
- `visualization`: The visualization object (if requested)

# Integration with Other tldrAI Features

## Using with tldr()

The script summarization feature complements the core `tldr()` function:

```{r}
# Get help with a specific R function
tldr("ggplot2::geom_boxplot")

# Summarize a script that uses this function
tldr_summarize("visualization.R")
```

While `tldr()` helps you understand specific R functions, `tldr_summarize()` helps you understand entire scripts and their structure.

## Using with Other tldrAI Functions

```{r}
# Generate code examples
tldr_example("how to create a boxplot faceted by two variables")

# Summarize the resulting script
tldr_summarize("boxplot_script.R")
```

This workflow can help you understand generated code or examples.

# Conclusion

The `tldr_summarize()` function provides a powerful way to quickly understand R scripts using AI-generated summaries and visualizations. By leveraging Large Language Models, it can identify the purpose, structure, and functionality of scripts without requiring you to read and analyze every line of code.

Whether you're exploring a new codebase, trying to understand old scripts, or documenting your work, `tldr_summarize()` can save you time and provide valuable insights into R code.

We recommend using this feature when:

- Onboarding to new projects with existing R code
- Revisiting old scripts you haven't worked with recently
- Trying to understand scripts written by colleagues
- Documenting your own scripts for future reference
- Teaching or explaining R code to others

The ability to focus on specific aspects of scripts and generate visualizations makes `tldr_summarize()` a versatile tool for R code comprehension and documentation.