---
title: "Using Visualizations in tldrAI"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Visualizations in tldrAI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Visualizing R Functions with tldrAI

One of the powerful features of `tldrAI` is its ability to generate visualizations that help understand R functions. These visualizations range from simple diagrams showing inputs and outputs to more complex representations of function relationships and code structure.

This vignette explains the visualization capabilities in detail and provides guidance on how to use them effectively.

## Visualization Types

`tldrAI` offers five different visualization types, each designed to highlight different aspects of R functions:

### 1. Diagram

The basic diagram shows a function's inputs, outputs, and package relationship in a simple flowchart:

```{r}
tldr("mean", visualize = TRUE, vis_type = "diagram")
```

This is useful for:
- Getting a quick overview of what a function does
- Understanding the function's inputs and outputs
- Seeing which package a function belongs to

### 2. Flowchart

The flowchart visualization shows the logical flow of a function, including conditional paths and loops:

```{r}
tldr("if", visualize = TRUE, vis_type = "flowchart")
```

This is useful for:
- Understanding control flow in functions
- Seeing how a function processes data
- Visualizing conditional logic and loop structures

### 3. Data Flow

The data flow visualization focuses on how data is transformed through a function:

```{r}
tldr("dplyr::filter", visualize = TRUE, vis_type = "data_flow")
```

This is useful for:
- Visualizing data transformation pipelines
- Understanding how data flows through dplyr, ggplot2, or stats functions
- Seeing the stages of data processing

### 4. Function Network

The function network visualization shows relationships between a function and other related functions:

```{r}
tldr("lm", visualize = TRUE, vis_type = "function_network")
```

This is useful for:
- Discovering related functions that work together
- Understanding function ecosystems within packages
- Finding alternative functions for similar tasks

### 5. Code Highlight

The code highlight visualization shows the actual function code with syntax highlighting:

```{r}
tldr("mean", visualize = TRUE, vis_type = "code_highlight")
```

This is useful for:
- Examining the implementation details of a function
- Learning from well-written R code
- Understanding exactly how a function works internally

## Enabling Visualizations

You can enable visualizations on a per-call basis:

```{r}
# Enable visualization for a specific call
tldr("mean", visualize = TRUE)
```

Or set visualizations to be enabled by default:

```{r}
# Enable visualizations by default
tldr_config(visualize = TRUE)
```

## Setting a Default Visualization Type

You can set your preferred visualization type:

```{r}
# Set default visualization type
tldr_visualization_config(enable = TRUE, default_type = "function_network")
```

Now all visualizations will use the function network type unless specified otherwise.

## Exporting Visualizations

You can export visualizations to files for use in documents, presentations, or sharing:

```{r}
# Export as SVG (vector graphics, good for presentations)
tldr("mean", visualize = TRUE, export_visualization = TRUE, 
     export_path = "mean_diagram.svg")

# Export as PNG (bitmap image, good for web)
tldr("lm", vis_type = "function_network", export_visualization = TRUE, 
     export_path = "lm_network.png", export_format = "png")

# Export as TXT (ASCII art, good for terminals/text documents)
tldr("if", vis_type = "flowchart", export_visualization = TRUE, 
     export_path = "if_flow.txt", export_format = "txt")
```

### Configuring Automatic Exports

You can configure visualizations to be automatically exported:

```{r}
tldr_visualization_config(
  auto_export = TRUE,               # Export all visualizations automatically
  export_format = "svg",            # Default export format
  export_dir = "~/R/visualizations" # Directory for exported files
)
```

With this configuration, every visualization will be automatically saved to the specified directory.

## Fallback Visualizations

If graphical packages (DiagrammeR, visNetwork, etc.) are not installed, `tldrAI` will automatically fall back to ASCII art visualizations:

```{r}
# ASCII diagram when DiagrammeR is not available
tldr("mean", visualize = TRUE)
```

These fallbacks ensure you still get valuable visualizations even in environments where installing graphical packages is not possible (like servers or restricted environments).

## Required Packages

Different visualization types require different R packages:

- **diagram**: Requires `DiagrammeR`
- **flowchart**: Requires `DiagrammeR`
- **data_flow**: Requires `DiagrammeR`, `htmlwidgets`
- **function_network**: Requires `visNetwork`, `igraph`
- **code_highlight**: Requires `htmltools`, `highlight`

For exporting:
- **SVG export**: Requires `DiagrammeRsvg` for diagrams
- **PNG export**: Requires `webshot` for all visualization types

You can configure whether to prompt for package installation:

```{r}
# Auto-install required packages without prompting
tldr_visualization_config(auto_install = TRUE)

# Get a visualization that requires packages to be installed
tldr("ggplot", vis_type = "function_network")
```

## Programmatic Access to Visualizations

You can work with visualizations programmatically:

```{r}
# Get function metadata
metadata <- get_function_metadata("mean")

# Create visualization object
vis <- create_visualization("mean", metadata, "diagram")

# Print visualization
print_visualization(vis)

# Export visualization
export_visualization(vis, "mean_diagram.svg")
export_visualization(vis, "mean_diagram.png", format = "png")
export_visualization(vis, "mean_diagram.txt", format = "txt")
```

## Examples for Different Function Types

Different visualization types are better suited for different kinds of functions:

### Data Manipulation Functions

For functions like `dplyr::filter`, `dplyr::mutate`, or `tidyr::pivot_longer`:

```{r}
# Data flow is best for data manipulation functions
tldr("dplyr::filter", vis_type = "data_flow")
```

### Statistical Functions

For functions like `lm`, `t.test`, or `anova`:

```{r}
# Function network is useful for statistical functions
tldr("lm", vis_type = "function_network")
```

### Control Flow Functions

For functions like `if`, `for`, or `while`:

```{r}
# Flowchart is best for control flow
tldr("if", vis_type = "flowchart")
```

### Base R Functions

For fundamental functions like `mean`, `sum`, or `apply`:

```{r}
# Code highlight is good for understanding base R
tldr("apply", vis_type = "code_highlight")
```

## Conclusion

Visualizations in `tldrAI` provide a powerful way to understand R functions from different perspectives. Whether you're learning a new function, exploring a package, or teaching others, these visualizations can enhance comprehension and make the learning process more engaging.

By using different visualization types for different functions and contexts, you can gain deeper insights into how R functions work and how they relate to each other within the broader R ecosystem.